\documentclass[]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

\newcommand{\Rey}{\ensuremath{Re}}
\newcommand{\ord}[1]{\ensuremath{\mathcal{O}(#1)}}
\newcommand{\degree}{\ensuremath{^{\circ}}}

\makeatletter
\newcommand{\rmnum}[1]{\romannumeral #1}
\newcommand{\Rmnum}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother

\newcommand{\modeone}{mode \Rmnum{1}}
\newcommand{\modetwo}{mode \Rmnum{2}}
\newcommand{\modethree}{mode \Rmnum{3}}
\newcommand{\modefour}{mode \Rmnum{4}}
\newcommand{\modefive}{mode \Rmnum{5}}
\newcommand{\modesix}{mode \Rmnum{6}}
\newcommand{\modeseven}{mode \Rmnum{7}}

\setlength{\textwidth}{180mm}
\setlength{\oddsidemargin}{-10mm}

\begin{document}

\section*{Responses to comments from reviewer 1}

We begin by thanking the reviewer for taking the time and effort
required to read and assess this manuscript. The changes we have made, along with explanations and
responses to the reviewer's comments, are outlined below. Any changes to the text related to the reviewer's comments are highlighted in
\textcolor{green}{green} Please note that typological errors pointed out were corrected but not highlighted.

\begin{itemize}
\item \emph{p.2	suppresses,	modern word, such as occurs in fluid-dynamic applications}

We thank the reviewer for the thorough reading required to make such suggestions. We have made the following changes to address these errors:
\begin{itemize}
\item ``suppresses'' changed to ``suppress''
\item ``word'' changed to ``world''
\item ``such as occurs in fluid-dynamic applications'' phrase changed to ``which is the setup in some fluid-dynamic applications''
\end{itemize}

\item \emph{p.3	systm,	performa	t high,	…at	low	Re,	Reare}

\begin{itemize}
\item ``systm'' changed to ``system''
\item ``Reare'' changed to ``Re are tested''
\end{itemize}

\item \emph{p.5	eq.1,(m),interpolating polynomials}

\begin{itemize}
\item ``(m)'' changed to ``m''
\item ``polynomials'' changed to polynomial
\end{itemize}

\item \emph{p.6	ad hoc…In fact forcing is obtained from	averaged lift force	in Joly et al. ?, the the}

The reviewer is right in stating that Joly et al deduce the magnitude of a forcing term from the averaged lift force from static cylinder simulations. However, it is stated in Joly et al  page 239 quoted (our emphasis)

``The characteristic angular frequency $\omega_{s}94$ is obtained from fixed cylinder simulations that provided closure coefficients of the QS model. To fit with FSI simulations, we take $F_0=0.66$ for the characteristic force. Note that we \emph{measured a lower value for its fixed simulation counterpart ($F_0=0.3$)}. This may reveal a more complex interaction with vortex shedding, which we would not deal with in this paper."

Therefore the value used as mean force $F_0= 0.66$ was chosen to provide the best match to the data, rather than from any theoretical consideration, and therefore can be considered as “ad hoc”.

\begin{itemize}
\item ``the the'' corrected to ``the''
\end{itemize}


\item \emph{p.7	is	a	proportional}

\begin{itemize}
\item ``is a proportional'' changed to ``is proportional''
\end{itemize}

\item \emph{p.10	Eq.	8	(m)}

\begin{itemize}
\item changed to ``m''
\end{itemize}

\item \emph{p.12 a a}

\begin{itemize}
\item changed to ``a''
\end{itemize}


\item \emph{why	unstable	branches	could	not	be	achieved	numerically	?	The	reviewer	thinks	
	it	is	possible	depending	on	how	you	perturb/perform	computations	?}

The reviewer is correct, it is technically possible to achieve unstable states numerically. The classic realisation of this is solving for steady states of a system using algorithms such as a Newton-Raphson method coupled with a continuation method to “follow” an unstable branch. The continuation method is required as the control parameter (in this case $\Pi_{2}$) cannot simply be incremented, but must be increased or decreased depending on the direction of the branch.

The situation here is more complex, in that the solutions are not steady states, but at their most simplest limit cycles (periodic solutions). Constructing a continuation method to follow periodic solutions is quite complicated. Considering the relatively minor improvement that the method would provide, we feel that the effort required is not justified.


\item \emph{p.	20 Fig.	8	caption	simulations	at.	…}

\begin{itemize}
\item ``.'' Removed 
\end{itemize}

\item \emph{p.22	$\Pi_{1}$ increases}

\begin{itemize}
\item space added between the two words
\end{itemize}

\item \emph{p.24	as	the	PI1	is…}

\begin{itemize}
\item removed ``the''
\end{itemize}


\end{itemize}

\end{document}