\chapter*{Acknowledgments}

I would like to express my deepest gratitude to my supervisors, Dr. Tan Boon Thong (Dr. Kenny) and Dr Justin Scott Leontini. Dr. Kenny thank you for entertaining me when I was having difficulties. Dr Justin, for explaining me complicated concepts and helping, guiding and providing words of encouragement when I was going through tough situations. I would like to thank both my supervisors for training me to conduct quality research and teaching me proper research techniques.

My warmest gratitude goes to the administrator of the Monash High-performance computer facility (SUN grid), Philip Chan for helping me out immensely and facilitating to carry out my simulations. Mention must be made to Monash University Malaysia for providing me with the scholarship to pursue my PhD. 
 
I would like to thank my closest friends Nishan and Hasuli who were pseudo-siblings for me from undergraduate level. Thank you for the support and encouragement. I would like to extend my thanks to Rangika and Devangi. Dr Anuja Dharmarathna, thank you for all the support, guidance and treating me as your own son when I needed help.

My Deepest gratitude goes out to my close friends in Monash Malaysia for helping me out, providing support and being there for me when I was going through very tough situations. 

I am indebted for the support provided by my family. Amma, Thatha, Akki, Ayya and Sanula (``Johnny"). Amma thank you for firmly telling me that `` A PhD has to be earned...!". A special thank goes out to Saranga for all the support you provided. My gratitude extends to Mr Nishantha Ranasinghe (Nishantha Ayya) for providing me strength and encouragement.
\\
\\
\\
Last but definitely not least, I am much indebted to Mrs Malin Bamunuarachchi for the blessings, prayers and the encouragement given to me when I hit ``rock-bottom"  and was on the verge of giving up. Thank you madam.    

 

  

  